---
output: html_document
---


```{r, echo=FALSE, message=FALSE}
# Summary information

# setwd("C:/Users/Soichi/Documents/info201/a6-mapping-shootings-Soichit")
# install.packages("plotly")
# install.packages("dplyr")
library(plotly)
library(dplyr)

# retrieve csv file data
shootings.df <- read.csv("data/Mass-Shooting-Data.csv")

num.shootings <- nrow(shootings.df)
num.injured <- sum(shootings.df$injured)
num.killed <- sum(shootings.df$killed)

# create summmary table info
summary.table <- c(num.shootings, num.injured, num.killed)
summary.table <- setNames(summary.table, c("Total Incidents", "Total Injuries", "Total Deaths"))


# group data by states to gather specific info
# this dataframe is used later for interactive plot2 and bonus map3
state.shootings <- shootings.df %>%
  group_by(state) %>%
  summarise( num_total = n(),
             injured_total = sum(injured),
             killed_total = sum(killed))
# adding state abbreviations column for choropleth map
abr <- state.abb[match(state.shootings$state,state.name)]
state.shootings <- mutate(state.shootings, abr)
#View(state.shootings)

# finding states that weren't affected by shootings
# data("state")
states.affected <- state.shootings$state
states.unaffected <- setdiff(state.name, states.affected)

# finding states with at least 10 incidents
states.high.incidents <- state.shootings %>% filter(num_total > 10) %>% select(-abr)
states.high.incidents <- setNames(states.high.incidents, c("State", "Total Incidents", "Total Injuries", "Total Deaths"))
```


# Mass shootings in the United States 2016 Data
<br />

### Summary
The data below shows mass shootings across the United States that occured in 2016.
**Data visualization** is used on this crowdsourced data with interactive charts to clearly 
display critical incident information in the locations they occured.
There were a total of **`r num.shootings`** incidents that occured this year. There were a
total number of **`r num.injured` injuries and `r num.killed` deaths** resulting from these
tragic incidents. 
<br />

##### Summary table
```{r, results='asis', echo=FALSE}
### creating table 1 ###
knitr::kable(summary.table)
```


```{r, echo=FALSE, message=FALSE}
# Description of a particular incident
worst.shooting <- filter(shootings.df, killed == max(killed))
# this data frame used for map and excludes outlier to fit all other markers within range
new.shootings.df <- filter(shootings.df, killed != max(killed))
# reformat date to look nice
shooting.date <- format(as.Date(worst.shooting$date, '%d-%B-%y'), format = '%B %d %Y')
```
<br />

### Incident at `r worst.shooting$city`, `r worst.shooting$state`
One particular incident stood out as one of the worst tragedies of 2016 in the United States.
This incident that occured on `r shooting.date` had more injuries and deaths than any other 
incidents that occured: **`r worst.shooting$injured` injuries and `r worst.shooting$killed` deaths**.
The shooting occured in a gay nightclub and is one of the **deadliest terroist attacks** in the 
United States since the September 11 attacks. Read more [here](https://en.wikipedia.org/wiki/2016_Orlando_nightclub_shooting).
<br /><br />


```{r, echo=FALSE, message=FALSE}
# An interactive map
# geo styling
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5
)
# create scatter on USA map
shootings.map <- plot_geo(new.shootings.df, lat = ~lat, lon = ~lng, colors=c("red","black")) %>%
  add_markers(
    text = ~paste(date, city, state, paste("Injuries:", injured), paste("Deaths:", killed), sep = "<br />"),
    color = ~injured,
    size = I(8),
    hoverinfo = "text"
  ) %>%
  colorbar(title = "# of Injuries") %>%
  layout(
    title = 'Mass Shootings Incidents across the United States', 
    geo = g
  )
```
### Interactive Map
`r shootings.map`
<br /><br />


```{r, echo=FALSE, message=FALSE}
# An interactive plot of your choice: Bar chart

# get bar plot data
bar.abr <- state.shootings$abr
bar.injured <- state.shootings$injured_total
bar.killed <- state.shootings$killed_total
df <- data.frame(bar.abr, bar.injured, bar.killed)

# create bar plot
shootings.bar.chart <- plot_ly(df, x = ~bar.abr, y = ~bar.injured, type = 'bar', name = 'Injuries') %>%
  add_trace(y = ~bar.killed, name = 'Deaths') %>%
  layout(title = 'Number of Injuries/Deaths by State',
         xaxis = list(title = 'States'),
         yaxis = list(title = 'Number of Injuries/Deaths'), barmode = 'group')

```
### Interactive Bar Chart
`r shootings.bar.chart`
<br /><br />

##### States with the most incidents
```{r, results='asis', echo=FALSE}
### creating table 1 ###
knitr::kable(states.high.incidents)
```
<br />


```{r, echo=FALSE, message=FALSE}
# An interactive plot 2 bonus
state.shootings$hover <- with(state.shootings, paste(state, '<br>', "Incidents:", num_total,
                             '<br>', "Injuries:", injured_total, "<br>", "Deaths:", killed_total))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)

# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa')
)

# create choropleth map
shootings.choropleth <- plot_geo(state.shootings, locationmode = 'USA-states',  colors=c("white","red")) %>%
  add_trace(
    z = ~num_total,
    text = ~hover,
    locations = ~abr,
    color = ~num_total
  ) %>%
  colorbar(title = "# of Incidents") %>%
  layout(
    title = 'States most affected by Mass Shooting Incidents',
    geo = g
  )
```
### Interactive Choropleth Map
`r shootings.choropleth`


##### States unaffected by any incidents (`r length(states.unaffected)` states)
```{r, results='asis', echo=FALSE}
### creating table 1 ###
knitr::kable(states.unaffected)
```


